```{r, message=FALSE}
library(tidyverse)
library(cytominer)
library(magrittr)
library(RCurl)
```

```{r, message=FALSE}
load_dataset  <- function(partition, dataset,feature){
    file_name  <- read_csv("../datasets.csv") 
    x  <-  file_name %>% filter(
         Partition == partition,
         Dataset == dataset,
         Features == feature) %>% 
         extract2("Link")

    return(read_csv(x) %>% 
          mutate(Metadata_dataset = dataset) %>%
          mutate(Metadata_partition = partition) %>% 
          mutate(Metadata_features = feature) 
          )
    }
```

# Load data 
We load training and test datasets for both genetic perturbation experiments 

```{r, message=FALSE}
# bbbc37 data 
bbbc037_train  <- load_dataset("Train","BBBC037","CellProfiler")  %>% 
    mutate(Metadata_x_mutation_status = "none")  %>%
  filter(str_detect(Metadata_pert_name, "WT") | Metadata_ASSAY_WELL_ROLE %in% c("Untreated", "CTRL")) 

bbbc037_test <- load_dataset("Test","BBBC037","CellProfiler")  %>% 
    mutate(Metadata_x_mutation_status = "none")  %>%
  filter(str_detect(Metadata_pert_name, "WT") | Metadata_ASSAY_WELL_ROLE %in% c("Untreated", "CTRL")) 

bbbc037 <- 
  bind_rows(bbbc037_train, bbbc037_test)
```

```{r, message=FALSE}
# bbbc043 data 
bbbc043_train  <- load_dataset("Train","BBBC043","CellProfiler")
 
bbbc043_test <- load_dataset("Test","BBBC043","CellProfiler")

bbbc043  <- bind_rows(bbbc043_train, bbbc043_test)
```

## Check dimensionality

```{r}
dim(bbbc043)
dim(bbbc037)
```

## Extract common features 

```{r}
colnames_bbbc037 <- colnames(bbbc037)
colnames_bbbc043 <- colnames(bbbc043)


Metadata_names_bbbc037 <- c(
   stringr::str_subset(colnames_bbbc037, "^Meta")
) 

Metadata_names_bbbc043 <- c(
   stringr::str_subset(colnames_bbbc043, "^Meta")
) 

common_metadata  <- intersect(Metadata_names_bbbc037, Metadata_names_bbbc043)  
common_features  <- setdiff(intersect(colnames_bbbc037, colnames_bbbc043),common_metadata)

```

# Concatenate data sets

```{r}
population  <- bind_rows(
    bbbc037 %>% 
        select(c(common_metadata, common_features)),
    bbbc043 %>% 
        select(c(common_metadata, common_features))
    ) %>% 
    mutate(Metadata_perturbation = "genetic") %>% 
    select(matches("^Meta"), everything())
```

## Important: update column names! 

```{r}
colnames_combined  <- colnames(population)

common_metadata  <- c(
   stringr::str_subset(colnames_combined, "^Meta")
) 

common_features  <- setdiff(colnames_combined, common_metadata)
```

Cytominer has problems handling column names '1', '2' so we rename them to 'Feature_1', ... 

```{r}
common_features  <- paste0("Feature_",common_features)
colnames(population)  <- c(common_metadata, common_features)
```

# Normalize data
We use cytominer to normalize both datasets with respect to the controls, i.e. EMPTY genes

```{r}
population_normalized  <- cytominer::normalize(
    population, 
    variables = common_features, 
    strata = c("Metadata_perturbation"), 
    sample = population %>% 
                filter(
                    Metadata_gene_name == 'EMPTY',
                    Metadata_partition == "Train"
                ), 
    operation = "standardize"
)
```

```{r}
population_normalized %>% dim() %>% print
```

# Aggregate data 

```{r}
population_aggregated  <- cytominer::aggregate(
    population = population_normalized, 
    variables = common_features, 
    strata = c("Metadata_gene_name","Metadata_dataset","Metadata_x_mutation_status"), 
    operation = "mean"
) 
```

```{r}
population_normalized %>% extract2("Metadata_gene_name") %>% print
```

```{r}
population_aggregated %>% slice(1:2) %>% print
```

# Correlation matrix 

```{r}
cor_matrix  <- cor(
    x = population_aggregated %>% 
        filter(Metadata_dataset == 'BBBC037') %>% 
        select(common_features) %>% 
        as.matrix() %>% 
        t, 
    y = population_aggregated %>% 
        filter(Metadata_dataset == 'BBBC043') %>% 
        select(common_features) %>% 
        as.matrix() %>% 
        t,
    use  = "complete.obs"
    ) 
```

# Submision file 

```{r}
# set column names 
colnames(cor_matrix)  <- population_aggregated %>% 
                            filter(Metadata_dataset == 'BBBC043') %>%
                            extract2("Metadata_x_mutation_status")

# set row names 
rownames(cor_matrix)  <- population_aggregated %>% 
                            filter(Metadata_dataset == 'BBBC037') %>%
                            extract2("Metadata_gene_name")


df  <- cor_matrix %>% as_data_frame() %>% 
            mutate(Metadata_gene_name = population_aggregated %>% 
                            filter(Metadata_dataset == 'BBBC037') %>%
                            extract2("Metadata_gene_name")) %>% 
            select(Metadata_gene_name, everything())

# write submission file
write.csv(df,"../cytodata-baseline_R.csv",row.names = FALSE)
```

```{r}
```